### 2018-04-05 Annotations/Annotations.tex
Added annotations

\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\begin{document}
\section{How to Generate a Good Word Embedding?}
Paper by Lai et al.\cite{lai2016generate}
\begin{itemize}
\item{Word embedding, also known as distributed word
representation, can capture both the semantic and syntactic
information of words from a large unlabeled corpus
and has attracted considerable attention from many researchers.
In recent years, several models have been proposed,
and they have yielded state-of-the-art results in many
natural language processing (NLP) tasks.}
\item{We observe that almost
all methods for training word embeddings are based on the
same distributional hypothesis: words that occur in similar
contexts tend to have similar meanings.}
\item{Training on a large corpus
generally improves the quality of word embeddings,
and training on an in-domain corpus can significantly
improve the quality of word embeddings for a specific
task.}
\item{Previous works have shown that models that predict the target word capture the paradigmatic relations between words}
\item{we can conclude that using a larger corpus can yield a better embedding, when the corpora are in the same domain}
\item{In most of the tasks, the influence of the corpus domain is
dominant. In different tasks, it impacts performance in the
different ways}
\item{The corpus
domain is more important than the corpus size. Using an indomain
corpus significantly improves the performance for a
given task, whereas using a corpus in an unsuitable domain
may decrease performance. }
\end{itemize}
\bibliographystyle{unsrt}
\bibliography{Bibliography}
\end{document}