### 2018-06-12 Thesis/Parts/Related_work/Related_work.tex

\documentclass[../../Thesis.tex]{subfiles}
\begin{document}
\header{Related work}
\end{document}