### 2018-06-14 Thesis/Parts/Motivation/Motivation.tex
Background, motivation and related work

\documentclass[../../Thesis.tex]{subfiles}
\begin{document}
\header{Motivation}
\subheader{In-domain embeddings and validation}
Earlier research, concerning domain specific articles, by \citet{Truong2017Thesis}, found that in-domain training of the word embeddings can  improve the process of document clustering. This effect is even stronger than the number of training examples and the model architecture. \citet{lai2016generate} found that the corpus domain is more important than the corpus size. Using an in-domain corpus significantly improves the performance for a given task, whereas using a corpus in an unsuitable domain may decrease performance.
Truong et al. encountered a problem in the validation of these in-domain embeddings. The word embedding produced correct document clustering results, leading to the conclusion that these embeddings are of good quality, since they capture the document relatedness needed to create correct clusterings. However unpublished results by Truong et al. state that the embeddings show high error rates on the validation scores. This seems to indicate that the word-vectors are of good quality, but that the available validation metrics cannot confirm this. \citet{Truong2017Thesis} used multiple word similarity validations to asses the quality of the word embeddings. However, these set are created to validate the generic embeddings, they fail to asses the quality of the domain specific embeddings.

\subheader{Research}
To asses the problem of the limited availability of pre-labelled validation sets for domain specific articles, we compare the embeddings to  TF-IDF on a categorization task. This A) gives an indication of the embedding quality for categorization tasks and B) contrasts the performance of embeddings to the performance of the more traditional TF-IDF approach. To ensure the quality of the embeddings for our research, we reuse the embeddings created in the research of \citet{Truong2017Thesis}.
\begin{jumpin}
	\textbf{RQ.1} Have word-embeddings a higher accuracy for academic texts than TF-IDF for article classification?\\
	\textbf{RQ.2} Which metric can be used to measure the accuracy of word embeddings for scientific articles?
\end{jumpin}
\subsubheader{Embedding and TF-IDF}
Research question one focusses on the classification results of both embedding-based techniques and TF-IDF.  To measure classification task we use the rank of the class to which the item belongs. Transforming the classification task from a binary metric to a ranking metric. For this task, we will use different versions of embeddings, and different TF-IDF versions to not only compare the two techniques, but also look for the optimal results of both techniques. For this part of the research, we will use the following hypothesis:
\begin{center}
\textit{Embedding based techniques give lower rankings than the TF-IDF based techniques}
\end{center}
By validating our invalidating  this hypothesis we get an indication of the performance of the embeddings compare to older techniques, and get insight into possible performance and resource trade-off's and if the computational power and AI expertise needed for the creation of embeddings is worth the investment .

\end{document}