### 2018-06-18 Thesis/Parts/Research_setup/Research_setup.tex

\documentclass[../../Thesis.tex]{subfiles}
\begin{document}

\header{Experiment setup}

\end{document}