### 2018-06-20 Thesis/Parts/Future_work/Future_work.tex

\documentclass[../../Thesis.tex]{subfiles}
\begin{document}
\header{Future work}
- Does the performance of embeddings and TF-IDF improve when embeddings and feature vectors are generated over the combination of title and abstract as one text?\\
	- this helps when dealing with articles that do not have a title or an abstract.\\
- How do unique token count, token count and vocabulary size relate for TF-IDF?\\
- Can high quality word similarity sets be created from embedding for future benchmarking?\\
- Can embeddings give a visual representation of the evolution of scientific research over time?\\
- Does TF-IDF improve when top-occurring words are cut-off?\\
\end{document}