### 2018-07-17 Thesis/Parts/Abstract/Abstract.tex

\documentclass[../../Thesis.tex]{subfiles}
\begin{document}
Over the last few years, word embeddings have taken a dominant position in the Information Retrieval domain. Many studies have been done concerning the quality and application of word embeddings on general texts, such as the Wikipedia corpus and comments on review websites. Giving promising results, the word embeddings have been studied and improved over recent years. However, these studies have been focused on generic texts, which are not limited to the characteristics of in-domain texts such as rare domain-specific words or have been focussed on small sets of academic texts. This research focusses on the quality and application of word embeddings on domain-specific texts, concerning a large corpus of 1.391.543 scientific articles which have been published in 2017. We validate the word embeddings using a categorization task to match articles to journals. We furthermore create a 2-dimensional visualization of the word embeddings on journal level, to visualize journal relatedness.\\
\vspace*{\fill}
\\
We aim to have this thesis published in the proceedings of the \textit{ECIR} (\url{http://ecir2019.org/}). \textit{Submission deadline: 9-10-2018}.
\end{document}