### 2018-07-20 Paper/paper.tex

\documentclass[runningheads]{llncs}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{subfiles}
\usepackage{amssymb}
\usepackage{csquotes}
\usepackage[numbers, sort, comma, square]{natbib}
\usepackage{changepage}
\usepackage{graphicx}
\usepackage{placeins}
\usepackage{xr}
\usepackage{tabularx}
\usepackage{lscape}
\usepackage{hyperref}
\usepackage{tablefootnote}
\usepackage{fancyhdr}
\usepackage{hyperref}
%\renewcommand\UrlFont{\color{blue}\rmfamily}

\newenvironment{jumpin}{\begin{adjustwidth}{0.3in}{}}{\end{adjustwidth}}

\begin{document}
\title{Document Embedding for Scientific Articles:\newline A validation of word embeddings}
\titlerunning{Document embedding for Scientific Articles}

\author{H.J. Meijer\inst{1,2}\orcidID{0000-1111-2222-3333} \and
R. Karimi\inst{2}\orcidID{1111-2222-3333-4444}}
\authorrunning{H.J. Meijer et al.}

\institute{University of Amsterdam, Science park 904, 1012WX Amsterdam, The Netherlands \and
Elsevier, Radarweg 29, 1043 NX Amsterdam, The Neterlands
\email{meijerarjan@live.nl, r.karimi@elsevier.com}}
\maketitle

\begin{abstract}
\subfile{Parts/Abstract/Abstract.tex}

\keywords{Word Embedding  \and Document Embedding \and Article Embedding \and Journal Embedding \and Embedding Visualization \and Embedding Validation.}
\end{abstract}

\subfile{Parts/Data/Data.tex}
%\subfile{Parts/Research_results/Research_results.tex}
%\subfile{Parts/Discussion/Discussion.tex}
%\subfile{Parts/Conclusion/Conclusion.tex}
%\subfile{Parts/Future_work/Future_work.tex}

%\begin{theorem}
%This is a sample theorem. The run-in heading is set in bold, while
%the following text appears in italics. Definitions, lemmas,
%propositions, and corollaries are styled the same way.
%\end{theorem}
%
% the environments 'definition', 'lemma', 'proposition', 'corollary',
% 'remark', and 'example' are defined in the LLNCS documentclass as well.
%
%\begin{proof}
%Proofs, examples, and remarks have the initial word in italics,
%while the following text appears in normal font.
%\end{proof}
\cite{Truong2017Thesis}
\bibliographystyle{splncs04}
\bibliography{Bibliography.bib}
\end{document}