### 2018-07-24 Paper/Parts/Future_work/Future_work.tex

\documentclass[../../Thesis.tex]{subfiles}
\begin{document}
\section{Future work}
\subsection{Intelligent cutting}
A better way of cutting could improve the quality of the embeddings. This improvement might be achieved by cutting the center of the vector space out before normalization. All words which are generic are in the center of the spectrum, removing these words prevents the larger texts to be pulled towards the middle of the vector space, where they lose the parts of their meaning which set them apart from the other texts. We expect that this way of cutting, instead of word-occurrence cutting, will improve the quality of the word embeddings.
\subsection{TF-IDFs performance point}
In our research, TF-IDF performed better on the abstracts than on the titles, which, according to us, is caused by the text size of the two texts. This leads to the question, how do token count and unique token count relate for TF-IDF? Is there a point at which the TF-IDF outperforms the embeddings, and will continue to outperform? If this relation is found, one could skip the TF-IDF calculations in certain situations, and skip the embedding training in other scenario's, reducing costs.
\subsection{Reversed word pairs}
At this point, there are no domain-specific word pair sets available. However, as we demonstrated, we can still test the quality of word embeddings. Once one has established that the word vectors are of high quality, could one create word pairs from the embeddings? If this is the case, we could create word pair sets using the embeddings, reverse-engineering the domain specific word pair sets for future use.
\end{document}