### 2018-07-25 Thesis/Parts/Abstract/Abstract.tex
update

\documentclass[../../Thesis.tex]{subfiles}
\begin{document}
Over the last few years, word embeddings have taken a dominant position in the information retrieval research. Studies conducted have focused on the quality and application of word embeddings on corpuses such as Wikipedia or online comments/reviews/tweets. However, these studies are limited to generic texts and lack technical, scientific or domain specific nuances such as rare words, lots of abbreviations, or chemical/mathematical formulas commonly used in academic context. This research focusses on the quality and application of word embeddings on academic corpuses. Word embedding requires less memory and is processed faster than conventional alternatives such as TFIDF. Hence, we intend to use word embeddings as an efficient way to model the content for our recommendation/disambiguation/search engines. This study focuses on word embedding compared to TFIDF for modeling content in scientific articles. We use a word2vec model  trained on titles and abstracts of roughly 68 millions scientific articles. Word embedding are evaluated against existing benchmarks for generic texts from previous studies. Furthermore, we have developed a new benchmark to evaluate these embedding in the scientific context. We validate the word embeddings via a categorization task to match articles to journals for roughly 1.4 millions articles published in 2017. Our results show that word embedding is a better content model for titles (short text) while TFIDF is a better model for abstracts (longer text). Furthermore, we have created a 2-dimensional visualization of the journals modeled via embeddings to visualize their relatedness. This graphs can be used to find competitive journals or gaps to propose new journals.\\
\vspace*{\fill}
\\
We aim to have this thesis published in the proceedings of the \textit{ECIR} (\url{http://ecir2019.org/}). \textit{Submission deadline: 9-10-2018}.
\end{document}